---
title: "Predict the manner in which they exercise using data from devices"
author: "Daniel Woo"
output:
  pdf_document: 
  html_document: default
    keep_md: yes
---

## Environment setup
```{r envsetup, echo=FALSE,message=FALSE}
library(caret)
library(mlbench)
```
## Executive Summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). We are particularly interested in the following two questions:

* The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with.

```{r loaddata, echo=FALSE, results='hide', message=FALSE}
setwd("C:/MOOC/module8/week3/project/Coursera-Practical-Machine-Learning-Project/data")

trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile)
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile)
}

trainData <- read.csv("./data/pml-training.csv")
testData <- read.csv("./data/pml-testing.csv")

dim(trainData)
dim(testData)
```

## Cleansing the dataset
```{r cleandata, echo=TRUE, results='hide'}

```

## Slicing the dataset
```{r slicingdata, echo=TRUE, results='hide'}
set.seed(7)
inTrain <- createDataPartition(y=trainData$classe, p=0.7, list=FALSE)
training <- trainData[inTrain, ]
testing <- trainData[-inTrain, ]

dim(training)
dim(testing)

training_prePCA <- training[,sapply(training, is.numeric)]
testing_prePCA <- testing[,sapply(testing, is.numeric)]

dim(training_prePCA)
dim(testing_prePCA)
```

## Preprocess the dataset
```{r preprocessdata, echo=TRUE, results='hide'}
training_preprocess <- preProcess(training_prePCA, method = "pca", thresh = 0.95)

print(training_preprocess)

training_PCA <- predict(training_preprocess, training_prePCA)
testing_PCA <- predict(training_preprocess, testing_prePCA)

dim(training_PCA)
dim(testing_PCA)

training <- cbind(training_PCA[,1:40], classe = training$classe)
testing <- cbind(testing_PCA[,1:40], classe = testing$classe)

dim(training)
dim(testing)
```

## Prepare training scheme
```{r preparetraindata, echo=TRUE, results='hide'}
control <- trainControl(method="repeatedcv", number=10, repeats=3)
```

Each model is automatically tuned and is evaluated using 3 repeats of 10-folds cross validation.

## Models
The three models constructed and tuned are Stochastic Gradient Boosting (also known as Gradient Boosted Machine or GBM), Random Forest (RF) and Recursive Partitioning (RPART).

## Train the GBM model

```{r gbmModel, echo=TRUE, results='hide', message=FALSE}
set.seed(7)
modGbm<-train(classe ~ ., method="gbm",data=training,verbose=FALSE)
```

## Train the RF model

```{r rfModel, echo=TRUE, results='hide', message=FALSE}
set.seed(7)
modRF<-train(classe ~ .,data=training,method="rf",prox=TRUE)
```

## Train the RPART model

```{r rpartModel, echo=TRUE, results='hide', message=FALSE}
set.seed(7)
modRpart <- train(classe ~ .,method="rpart",data=training)
```

## Collect the resamples
results <- resamples(list(GBM=modGbm, RF=modRF, RPART=modRpart))

## Summarize the distributions
summary(results)

## Predicting for test data sets
qplot(predict(modGbm,testing),trainData,data=testing)
qplot(predict(modRF,testing),trainData,data=testing)
qplot(predict(modRpart,testing),trainData,data=testing)


## Appendix: Figures

### boxplots of results
bwplot(results)

### dot plots of results
dotplot(results)
